\chapter{Final Remarks}
\label{cha:final_remarks}

\section{Discussion}
This thesis explained and compared two authentication mechanisms for the service-to-service security in a microservice deployment.
Both mechanisms are based on public key cryptography, but each mechanism has its own advantages and disadvantages.

Mutual TLS is a very efficient and straightforward authentication mechanism.
The implementation of mTLS is very simple, since the work is handled by the TLS protocol.
mTLS does not provide many configuration parameters and it does not allow to add custom functionalities, like sharing the user context without another technology.
Nevertheless if an developer aims to implement only service-to-service authentication, mTLS is the superior authentication mechanism.

As soon as nonrepudiation is an requirement self signd JWTs are the superior authentication mechanism.
Furthermore JWTs make the identity propagation more convenient and allow the developers to customize the authentication mechanism and add additional parameters.
On the other hand the implementation of self-signed JWTs is more challenging and requires each service to know how to work with JWTs.
Therefore choosing JWTs over mTLS would be unnecessary overhead when the target is implementing only service-to-service authentication.
The decision if the additional control of the approach using self-signed JWTs is worth the overhead has to be evaluated for each project independently.

The biggest challenge of both authentication mechanisms is the key-management.
Both mechanisms require a self hosted PKI and require to handle all associated tasks.
Therefore the implementation of the authentication mechanisms is much less work than the key management.
Nevertheless, the level of security that is provided using public key cryptography, is worth the expenses.

It is very important to mention that the requirement of service-to-service authentication is only a consequence of the microservice architecture.
Before evaluating the different authentication mechanism, developers should evaluate if the benefits of the microservice architecture are worth the resulting consequences.

\section{Summary}
