\chapter{Final Remarks}
\label{cha:final_remarks}

\section{Discussion}
This thesis compared and explained the most popular mechanisms for mutual authentication in a microservice deployment.
Both mechanisms are based on public-key cryptography, but each mechanism has its own motivations and challenges.

Mutual TLS is an efficient and straightforward authentication mechanism.
The implementation of mTLS is very simple since the TLS protocol defines the processes.
Otherwise, mTLS does not provide many configuration parameters, and it does not allow to add custom functionalities, like sharing the user context without additional technologies.
Nevertheless, if a developer aims to implement only service-to-service authentication, mTLS is the preferred authentication mechanism.

As soon as nonrepudiation is required, self-signed JWTs are the superior authentication mechanism.
Furthermore, JWTs make identity propagation more convenient and allow the developers to customize the authentication mechanism and add additional parameters.
On the other hand, the implementation of self-signed JWTs is more challenging and requires each service to know how to work with JWTs.
Therefore choosing JWTs over mTLS would be unnecessary overhead when the target is implementing only service-to-service authentication.
The decision if the additional control of the approach using self-signed JWTs is worth the overhead has to be evaluated for each project independently.

The experiment of chapter~\ref{cha:experiment} showed that the performance of the compared authentication mechanisms is very similar.
According to the experiment results, the performance of the approach using self-signed JWTs is very dependent on the implementation of the mechanisms.
In some cases, using self-signed JWTs results in a lower request duration, but in some situations, it results in a higher request duration.
Therefore the performance is not a criterion that makes any mechanism superior to the other.

The biggest challenge of both authentication mechanisms is key management.
Both mechanisms require a PKI and require handling all associated key management tasks.
Therefore, implementing the authentication mechanisms is less challenging than key management.
Nevertheless, the level of security that is provided using public-key cryptography is worth the expenses.

\section{Summary}
Service-to-service authentication is a requirement caused by the migration to the microservice architecture.
Function calls within the monolithic backend migrate to remote calls.
Remote calls have to assure authentication, confidentiality, and integrity.
Confidentiality and integrity can be assured using TLS, but authentication of both parties requires additional mechanisms.

This thesis compared two of the most popular authentication mechanisms for service-to-service authentication.
Therefore the fundamentals and concepts of the compared authentication mechanisms were described in detail.
Additionally, a project using the discussed mechanisms was reviewed, and the consequences of the different mechanisms were described.
In the end, an experiment comparing the performance of the discussed authentication mechanisms was performed, and the results were discussed.

